Red, Blue, or Purple? 2024’s Battleground States Tell a Story

### 2024-11-04 paper/paper.qmd
paper final draft

package_list <- c("tidyverse", "arrow", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "writexl", "gridExtra", "grid", "usmap")

@@

Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

The final predictions for the 2024 presidential election indicate a close race between the two candidates. The model results are summarized in @tbl-modelresults. Based on the weighted average calculations, which incorporate factors such as state effects, pollster reliability, candidate-specific adjustments, and recent polling trends, Donald Trump is predicted to receive approximately 46.56% of the vote, while Kamala Harris is predicted to receive around 45.65%. These predictions account for adjustments based on sample size, recency of the polls, and reliability scores, ensuring that more recent and reliable polls have a greater influence on the final outcome. The model also applies a slight adjustment to account for known biases, providing a robust estimation of each candidate’s standing. Although these predictions offer a snapshot of the current polling landscape, they should be interpreted with caution, as last-minute changes in voter sentiment could still impact the final outcome.

@@

# Add a bias for Trump
trump_bias <- 0.95 
data <- data %>%
  mutate(
    y_pred = ifelse(candidate_name == "Donald Trump", y_pred + trump_bias, y_pred)
  )

>>>>

package_list <- c("tidyverse", "arrow", "kableExtra", "ggplot2", "dplyr", "here", "knitr", "rstanarm", "modelsummary", "readr", "gridExtra", "grid", "usmap")

@@

The model was run in R [@citeR] using the `rstanarm` package [@citeRstanarm] to implement a Hierarchical Bayesian Model with random intercepts for states and pollsters. Weakly informative priors were specified, including a normal prior with mean 0 and standard deviation 2.5 for both the coefficients and intercept, and an exponential prior with a rate of 1 for the auxiliary (residual variance) parameter, in line with the default settings in `rstanarm`. These priors serve to regularize the estimates, preventing extreme values while allowing the data to guide the posterior distributions.

Background details and diagnostics are included in [Appendix @sec-model-details].

@@

The final predictions for the 2024 presidential election indicate a close race between the two candidates. The model results are summarized in @tbl-modelresults. Based on the weighted average calculations, which incorporate factors such as state effects, pollster reliability, candidate-specific adjustments, and recent polling trends, Donald Trump is predicted to receive approximately 47 of the vote, while Kamala Harris is predicted to receive around 45%. These predictions account for adjustments based on sample size, recency of the polls, and reliability scores, ensuring that more recent and reliable polls have a greater influence on the final outcome. The model also applies a slight adjustment to account for known biases, providing a robust estimation of each candidate’s standing. Although these predictions offer a snapshot of the current polling landscape, they should be interpreted with caution, as last-minute changes in voter sentiment could still impact the final outcome.

@@

# Add biases for Trump and Harris 
trump_bias <- 12.4 
harris_bias <- 10.21

data <- data %>%
  mutate(
    y_pred = case_when(
      candidate_name == "Donald Trump" ~ y_pred + trump_bias,
      candidate_name == "Kamala Harris" ~ y_pred + harris_bias,
      TRUE ~ y_pred
    )
  )
